
\newglossaryentry{gl:lstm}{
	name={Long Short Term-Memory},
	description={A \gls{gl:rnn} with a memory cell to distribute information along the other RNN cells.}
}\newacronym{ac:lstm}{LSTM}{Long Short Term-Memory}

\newglossaryentry{gl:rnn}{
	name={Recurent Neural Network},
	description={A Neural network that passes information between cells in the same layers.}
}

\newglossaryentry{gl:hashmap}{
	name=Hashmap,
	description={A list of items where their unique placmnt in the list is detemend by their unique refrence key using a function that maps the key to a placement in the list.}
}

\newglossaryentry{gl:arima}{
	name={Autoregressive Integrated Moving Average},
	description={A form of regression model that is based on previus timesteps. This model uses 3 models; Autoregressive terms, Itegration terms, and moving average to }
}\newacronym{ac:arima}{ARIMA}{\gls{gl:arima}}

\newglossaryentry{gl:dataframe}{
	name=DataFrame,
	description={A table of values. The name is from the python library Pandas used in this study.}
}

\newglossaryentry{gl:hype_param}{
	name=Hyper-Parameter,
	description={A parameter in a deep-learning model that adjustes the learning prosess.}
}

\newglossaryentry{gl:stl}{
	name={Seasonal-Trend decomposition using LOESS},
	description={Takes a timeseries and decomposing it into a trend component, season component, and residual component using local regression for smoothing\cite{cleveland_stl_1990}.}
}\newacronym{ac:stl}{STL}{\gls{gl:stl}}

\newglossaryentry{gl:mstl}{
	name={Multiple Seasonal-Trend decomposition using LOESS},
	description={Based in the traditional \gls{gl:stl} it decomposes a timeseries into several seasons, trend, and residual\cite{bandara_mstl_2021}.}
}\newacronym{ac:mstl}{MSTL}{\gls{gl:mstl}}

\newglossaryentry{gl:masl}{
	name={Meters above sea level},
	description={The altitude above mean sea level.}
}\newacronym{ac:masl}{m.a.s.l}{Meters above sea level}

% Independent accronyms

\newacronym{ac:nibio}{Nibio}{Norwegian Institute of Bioeconomy Research Agrometeorology service}
\newacronym{ac:kilden}{LMT}{\acrfull{ac:nibio}}
\newacronym{ac:nibio_n}{Nibio}{Landbruksmeteorologisk Tjeneste}
\newacronym{ac:kilden_n}{LMT}{\acrfull{ac:nibio_n}}
\newacronym{ac:met}{MET}{The Norwegian Meteorological Institute}
\newacronym{ac:nmbu}{NMBU}{The Norwegian University of Life Science}
\newacronym{ac:rmse}{RMSE}{Root Mean Square Error}
\newacronym{ac:mae}{MAE}{Mean Absolut Error}
\newacronym{ac:r2}{$R^2$}{Explained Variance}
\newacronym{ac:kappa}{$\log(\kappa)$}{Log Condition number}
\newacronym{ac:gru}{GRU}{Gated Recurent Unit}
\newacronym{ac:fdm}{FDM}{Finite Difference Method}
\newacronym{ac:asper}{ASPER}{Answer Set Programming-enhanced Entity-Relation}